\LARGE \bf Approximately Orchestrated Road Traffic Automata\\ a General Traffic Simulation Framework

### 2012-03-14 papers/itsc_2012/final.tex
splitting and reorganizing simulation section

AORTA's implementation uses Scala, a language implemented on the Java Virtual
Machine \cite{scala}. Scala provides the advantage of functional programming
constructs while still permitting imperative style.  Extensions and clients
built on top of AORTA can be written in either Java or Scala. The software is
open-source \footnote{source code at http://code.google.com/p/road-rage} and
easily extensible. For instance, the stop sign sign policy used in experiments
described in section \ref{sec:result} is implemented in just 70 lines of code.

@@

AORTA uses microscopic simulation, modeling individual drivers as agents.
Simulation is discrete-time, meaning an agent accelerates at a fixed rate for
the duration of a ``step'' to achieve a new position and velocity, and
space-continuous, meaning roads are not tiled \commentpk{recommend breaking up
preceding sentence into 2. I lost track haflway}. The step duration dt is
fixed, enabling agents to reason about collision avoidance and satisfying other
constraints \commentpk{preceding sentence is unclear}. Setting this value to
the average ``reaction time'' of a human driver may be appropriate for some
\commentpk{"human driver" instead of some?} experiments. Each step, the
simulation does the following:

\begin{enumerate}
  \item Introduces new agents into the map
  \item Updates the position and velocity of agents based on their choice of
  acceleration in the previous step
  \item Checks for collisions
  \item Allows each agent to choose a reaction \commentpk {should be action} for the next step
\end{enumerate}

Agents have a few enforced constraints -- they cannot exceed their physical
acceleration capabilities, travel in reverse, or travel past the end of a
traversable without appearing on the next. Intersection policies and avoiding
collisions on the same traversable provide the other constraints, but the
agent's behavior is responsible for obeying these \commentpk{if these "other"
constraints are not enforced, clearly specify. Something like "Agent behaviors
are responsible for following intersection policies as directed by an
intersection, and are responsible for not colliding with other agents" probably
sounds a bit better}. There are three main components of the simulation, each
configurable by extending built-in implementations of an interface: agent
behaviors, routing strategies, and intersection policies.

Currently agents cannot switch lanes. AORTA allows merging lanes by treating
the merge point as an intersection. \commentpk{i take it this is an orphan
paragraph?}

\commentpk{this section is a bit incoherent, and might need to be reorganized.
should the constraints be mentioned here explicitly, or maybe just added to
agent behaviors? What are you introducing in the next few subsections? I can
see A-E working together as an explanation of the workings of the simulator,
and would make more sense right after the enumeration above. Section G can be
moved into the experiments section. Once you are done with this primary
explanation, Section F should be talked about separately under something along
the lines of current limitations of the simulator.}

@@

To introduce new agents into the simulation in real-time, users can create
\emph{generators} programatically or using the UI. A generator picks a random
starting position and goal from two sets of roads \commentpk{preceding sentence
is ambiguous, maybe go with "The generator produces a desired amount of traffic
by initializing agents from random locations inside a \textit{start} polygon to
random destinations inside a \textit{goal} polygon, as demarcated by the
user"}. Users can select these roads by mouse with a bounding polygon in the
UI, allowing specific traffic patterns like rush-hour scenarios to be easily
simulated.  A generator \commentpk{polygon} may also cover the entire map,
uniformly distributing traffic everywhere.

Every step, a generator creates some number of new agents. If the agent's route
policy needs computation like pathfinding to operate, the generator either
performs it immediately (incurring a delay in simulation if it is running) or
sends it to a pool of worker threads to process in the background \commentpk{if multiple threads are available?}.

Once an agent's route is ready, the agent waits alongside its starting lane (as
if it is in a driveway or parking lot). When the lane has no agents that could
potentially crash with the new vehicle, the new agent enters the system with
zero initial speed.  Because behaviors must enforce speed limit along a lane
from the moment an agent enters a road, it is safe for a new agent to enter a
road if there is no other agent within the worst-case distance (traveling
distance at the speed limit, plus stopping distance from this speed) preceeding
the spawn point.  To avoid predicting which agents might enter the lane in the
near future, there is a buffer space at the beginning of the lane where a new
agent may never spawn. This disqualifies some short lanes from being spawn
points, but this is not a big limitation in practice. \commentpk{I understand
your concern about explaining the entire methodology, but the last 2 lines are
unnecessary and somewhat distracting}

@@

\subsection{Agent Reactions}

Each agent has a \emph{behavior} governing it. When the agent travels past the end
of a lane during its step, the behavior picks the turn to take. At each step,
the behavior can make the agent react by performing one of two actions:
disappearing from the map (when the agent is at rest and is done with its route)
or setting an acceleration for the next step. Were lane changing modeled, that
would be another possible action.

@@

\commentd{one diagram probably covers lots of this}

@@

@@

\subsection{Gridlock}

\subsection{Determinism and scale}

AORTA is designed for controlled experiments. Consequently, simulations are
deterministically reproducible. A seed for the pseudo-random number generator,
the input graph, and the generators' configuration fully determine the outcome of any
simulation. This determinism may break down when generators use worker threads,
since routes may be computed in different orders, causing agents to enter the
road at different times.

\commentd{saving a ``scenario'' can soon be a feature}

\commentd{and how will we solve this issue with threads and determinism?}

\commentd{if i have a day after the paper's done, I can make this number
better.}

@@

@@

The number of agents changes because the continuous generator introduces a new
agent every second. This count reaches a steady state when the generator
introduces an agent at the same rate as older agents finish their route. Figure
\ref{agent_cnt} reveals that autonomous reservations let agents reach their
destination twice as quickly as the competition, with a steady-state of about
400 agents. The count for traffic signals continues to increase because
gridlock occurred in one segment of the map, causing agents to not finish their
route. Unexpected starvation was also observed: although the signals cycle
through turns regularly, agents cannot proceed when the previous cycle caused a
destination lane to fill and become blocked.

>>>>

AORTA's implementation uses Scala, a language implemented on the Java Virtual
Machine \cite{scala}. Scala provides the advantage of functional programming
constructs while still permitting imperative style.  Extensions and clients
built on top of AORTA can be written in either Java or Scala. The software is
open-source \footnote{source code at http://code.google.com/p/road-rage} and
easily extensible. For instance, the stop sign sign policy used in experiments
described in section \ref{sec:results} is implemented in just 70 lines of code.

@@

AORTA uses microscopic simulation, modeling individual drivers as agents.
Time is modeled discretely, meaning an agent accelerates at a fixed rate for
the duration of a ``step'' to achieve a new position and velocity. Space is
continuous, meaning agents occupy a moving interval of a lane, rather than one
fixed tile of the road. Each step lasts the same fixed duration \textit{dt}.
This lets agents reason about how far they will travel for a given choice of
acceleration before being able to choose a different acceleration. Setting
\textit{dt} to the ``reaction time'' of a human driver may be appropriate for
experiments modeling humans.

Currently agents cannot switch lanes. AORTA allows merging lanes by treating
the merge point as an intersection. \commentpk{i take it this is an orphan
paragraph?} \commentd{yeah, where should it belong?}

Each step, the simulation does the following:

\begin{enumerate}
  \item Introduces new agents into the map
  \item Updates the position and velocity of agents based on their choice of
        acceleration in the previous step
  \item Checks for collisions
  \item Allows each agent to choose an action for the next step
\end{enumerate}

Each step (except 2) will be described below, followed by limitations of the
simulator. Section \ref{sec:config} will describe how researchers may easily
modify relevant portions of AORTA.

@@

To introduce new agents into the simulation in real-time, users can create
\emph{generators} programatically or using the UI. The generator produces the
desired amount of traffic by initializing agents from random locations inside a
\textit{start} polygon to random destinations within a \textit{goal} polygon.
Users can draw these polyons by mouse in the UI, allowing specific traffic
patterns like rush-hour scenarios to be easily simulated. A generator polygon
may also cover the entire map, uniformly distributing traffic everywhere.

Every step, a generator creates some number of new agents. If the agent's route
policy needs computation like pathfinding to operate, the generator either
performs it immediately (incurring a delay in simulation if it is running) or,
if possible, sends it to a pool of worker threads to process in the background.

Once an agent's route is ready, the agent waits alongside its starting lane (as
if it is in a driveway or parking lot). When the lane has no agents that could
potentially crash with the new vehicle, the new agent enters the system with
zero initial speed. Because behaviors must enforce speed limit along a lane
from the moment an agent enters a road, it is safe for a new agent to enter a
road if there is no other agent within the worst-case distance (traveling
distance at the speed limit, plus stopping distance from this speed) preceeding
the spawn point.% To avoid predicting which agents might enter the lane in the
%near future, there is a buffer space at the beginning of the lane where a new
%agent may never spawn. This disqualifies some short lanes from being spawn
%points, but this is not a big limitation in practice. \commentpk{I understand
%your concern about explaining the entire methodology, but the last 2 lines are
%unnecessary and somewhat distracting}

@@

\subsection{Agent Reactions}
\label{sec:behaviors}

Each agent has a \emph{behavior} governing it, responsible for obeying
intersection policies and avoiding collisions. When the agent travels past the
end of a lane during its step, the behavior picks the turn to take. At each
step, the behavior can make the agent react by performing one of two actions:
disappearing from the map (when the agent is at rest and is done with its route)
or setting an acceleration for the next step. Were lane changing modeled, that
would be another possible action.

@@

@@

\subsection{Simulation Limitations: Gridlock}

\section{CONFIGURABILITY}
\label{sec:config}

There are three main components of the simulation, each configurable by
extending built-in implementations of an interface: agent behaviors, routing
strategies, and intersection policies. Behaviors have been described already in
section \ref{sec:behaviors}.

@@

@@

Each experiment is deterministically reproducible. A seed for the pseudo-random
number generator, the input graph, and the generators' configuration fully
determine the outcome of any simulation \footnote{This determinism may break
  down when generators use worker threads, since routes may be computed in
different orders, causing agents to enter the road at different times.}. After
setting up a scenario in the UI, users can save this configuration for
resimulation.

@@

The number of agents changes because the continuous generator introduces a new
agent every second. This count reaches a steady state when the generator
introduces an agent at the same rate as older agents finish their route. Figure
\ref{fig:agent_cnt} reveals that autonomous reservations let agents reach their
destination twice as quickly as the competition, with a steady-state of about
400 agents. The count for traffic signals continues to increase because gridlock
occurred in one segment of the map, causing agents to not finish their route.
Unexpected starvation was also observed: although the signals cycle through
turns regularly, agents cannot proceed when the previous cycle caused a
destination lane to fill and become blocked.
